Semi-permeable vesicles: mathematical formulation, numerics and physics

### 2020-05-04 vesicle_code/docs/permeableVes/main.tex
working on relaxation

\title{Semi-permeable vesicles: mathematical formulation, numerics and physics}
\author{Ashley Gannon, Bryan Quaife, Shravan Veerapaneni, and Yuan-Nan Young}

@@

\section{Problem formulation} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{subequations}
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  &&\xx \in \RR^2 \setminus \gamma, \\
% 
  \nabla \cdot \uu &= 0,  &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    \qquad &&\norm{\xx} \rightarrow \infty.
\end{alignat}
\end{subequations}%
%
Letting $\nn$ be the outward normal to $\gamma$, $\jump{\cdot}$ denote
the jump across the interface, and $T$ denote the hydrodynamic stress
tensor, the boundary conditions along the vesicle are
\begin{align}
  \label{eqn:jump1}
  \jump{T\cdot\nn} &= \ff_\text{mem}, \\
  \label{eqn:jump2}
  \jump{\uu} &= 0,
\end{align}
where $\nn$ points outward to the vesicle, and $\ff_\text{mem}$ is the
standard membrane force that includes a bending and a tension term. In
particular, $\ff_\text{mem} = \ff_\text{ben} + \ff_\text{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\text{ben} = -\xx_{ssss}, \qquad
  \ff_\text{ten} = (\sigma \xx_s)_s.
\end{align}
Since the bending stiffness only sets the time scale, we let $k_b = 1$
throughout this paper. The Lagrange multiplier $\sigma$ is used to
satisfy the inextensibility condition.  Finally, the tension of the
vesicle is
\begin{align}
  \Lambda = \sigma + 1.5 \kappa^2,
\end{align}
where $\kappa$ is the membrane curvature.

The kinematic condition is modified to account for the fact that the
fluid can move through the membrane. A standard choice appears
is~\cite{}
\begin{align}
  \uu - \dot{\xx} = - \beta (\ff_\text{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is a semi-permeability rate.

\subsection{Integral equation formulation}
Since the flow is governed by the linear Stokes equations, it can be
represented in terms of layer potentials. Given the boundary
conditions~\eqref{eqn:jump1} and~\eqref{eqn:jump2} required on the
interface, the fluid velocity can be represented as the single-layer
potential
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\text{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
where $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Then, by the
kinematic equation, the dynamics of the vesicle 
\begin{align}
  \label{eqn:vesVelocity}
  \frac{d\xx}{dt} = \uu_\infty(\xx) + \beta (\ff_\text{mem}\cdot\nn)\nn
  + \SS [\ff_\text{mem}](\xx).
\end{align}
Finally, the inextensibility condition is given by
\begin{align}
  \xx_s\cdot\dot{\xx}_s = 0.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Vesicle Area}
Since the vesicle membrane is semi-permeable, the vesicle's area
fluctuates.  Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
\end{align}
The time derivative of $A(t)$ includes terms due to the time derivative
of $\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, this second term is zero
because of the local inextensibility condition. Therefore, 
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality comes follows from the divergence theorem.
Finally, using integration by parts and the Frenet-Serret identities, we
have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= \beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= \beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= \beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= \beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align}
Finally, using the relationship between the Lagrange multiplier
$\sigma$, and the tension $\Lambda$, the area of a semi-permeable
vesicle is governed by
\begin{align}
  \dot{A}(t) = \beta \int_\gamma \left(
    -\frac{\kappa^3}{2} + \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
We consider a single semi-permeable vesicle in four different flow
regimes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow}
\begin{itemize}
  \item Always goes to a circle
  \item Area formula works
\end{itemize}

We consider a single semi-permeable vesicle in a quiescent flow. For
impermeable vesicles, the vesicle relaxes to a shape where the tension
$\Lambda$ is constant and negative~\cite{kra-win-sei-lip1996}. In the
semi-permeable case, the area of the vesicle is governed by
equation~\eqref{eqn:areaROC}, where $\beta > 0$. Therefore, when the
semi-permeable vesicle reaches a steady-state shape, we have
\begin{align}
  \int_{\gamma} \left(-\frac{\kappa^3}{2} + \kappa \Lambda \right) \, ds
  = 0.
\end{align}
Based on numerical observations, the steady-state shape is a circular
vesicle with constant tension. Since the tension and curvature are
constant, we can solve for the curvature in terms of the tension $\kappa
= \sqrt{2\Lambda}$. Therefore, the steady state shape is a circle of
radius $r = 1/\sqrt{2\Lambda}$, and therefore the length is
\begin{align}
  L = \frac{\sqrt{2}\pi}{\sqrt{\Lambda}},
\end{align}
which is constant by the local inextensibility constraint. Finally, we
can compute the final tension in terms of the length using the formula
\begin{align}
  \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:finalTension}
\end{align}
Therefore, in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

We start by simulating a star-shaped vesicle in a quiescent flow with
semi-permeability rate $\beta=1$ (Figure~\ref{fig:starShape}). To
perform this simulation, we use an adaptive second-order time stepping
method~\cite{qua-bir2016}. The times and reduced areas are in the
titles, and note that the shapes are not eqiuspaced in time.
\begin{figure}[htp]
  \begin{minipage}{0.55\textwidth}
    \input{figures/starRelaxation1.tikz}
    \input{figures/starRelaxation2.tikz}
    \input{figures/starRelaxation3.tikz}
    \input{figures/starRelaxation4.tikz}
    \input{figures/starRelaxation5.tikz}
    \input{figures/starRelaxation6.tikz}
  \end{minipage}
  \hfill
  \begin{minipage}{0.4\textwidth}
  \input{figures/starAreaDeriv.tikz}
  \end{minipage}
  \caption{\label{fig:starShape} A initially star-shaped semi-permeable
  vesicle in a shear quiescent flow with $\beta=1$. Note that the time
  steps are not equispaced. The area (red) and its derivative (blue) of
  an initially star-shaped vesicle as a function of time. The dots
  correspond to the time steps shown in six time steps
  show on the left.}
\end{figure}

\begin{figure}[htp]
\begin{minipage}{0.40\textwidth}
\input{figures/starTension.tikz}
\end{minipage}
\hfill
\begin{minipage}{0.55\textwidth}
  \centering
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime1.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime2.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime3.pdf} \\
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime4.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime5.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime6.pdf}
\end{minipage}
  \caption{\label{fig:starTension} The tension of the time steps shown
  in Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$
  is also included. At this point, the vesicle is nearly circular and
  the tension is constant and positive. The constant is given by
  equation~\eqref{eqn:finalTension}, and this value corresponds to the
  dashed black line.}
\end{figure}

\begin{figure}[htp]
\begin{minipage}{0.40\textwidth}
\input{figures/starFlux.tikz}
\end{minipage}
\hfill
\begin{minipage}{0.55\textwidth}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime1.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime2.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime3.pdf} \\
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime4.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime5.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime6.pdf}
\end{minipage}
  \caption{\label{fig:starFlux} The flux of the time steps shown in
  Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$ is
  also included. At this point, the vesicle is nearly circular and there
  is no flux.}
\end{figure}

@@

%\ifTikz
\input{figures/shearPhaseDiagram.tikz} 
%\fi

@@

>>>>

\title{Semi-permeable vesicles: mathematical formulation, numerics and physics}
\author{Ashley Gannon, Bryan Quaife, and Yuan-Nan Young}

@@

\section{Problem formulation} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{subequations}
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  &&\xx \in \RR^2 \setminus \gamma, \\
% 
  \nabla \cdot \uu &= 0,  &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    \qquad &&\norm{\xx} \rightarrow \infty.
\end{alignat}
\end{subequations}%
%
Letting $\nn$ be the outward normal to $\gamma$, $\jump{\cdot}$ denote
the jump across the interface, and $T$ denote the hydrodynamic stress
tensor, the boundary conditions along the vesicle are
\begin{align}
  \label{eqn:jump1}
  \jump{T\cdot\nn} &= \ff_\{mathrm{mem}, \\
  \label{eqn:jump2}
  \jump{\uu} &= 0,
\end{align}
where $\nn$ points outward to the vesicle, and $\ff_\mathrm{mem}$ is the
membrane force that includes a bending term that minimizes the Helfrich
energy, and a tension term that acts as a Lagrange multiplier to satisfy
the local inextensibility condition. The bending stiffness only sets the
time scale, so we let $k_b = 1$ throughout the paper. For ease of
implementation, we modify the bending tension and forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} + \ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -\xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
However, the physical tension $\Lambda$ can be computed as a
post-processing step as
\begin{align}
  \Lambda = \sigma + 1.5 \kappa^2,
\end{align}
where $\kappa$ is the membrane curvature.

The kinematic condition is modified to account for the fact that the
fluid can move through the membrane. A standard choice appears
is~\cite{}
\begin{align}
  \uu - \dot{\xx} = - \beta (\ff_\mathrm{mem} \cdot \nn) \nn, \qquad
  \xx \in \gamma,
\end{align}
where $\beta$ is a semi-permeability rate.

\subsection{Integral equation formulation}
Since the flow is governed by the linear Stokes equations, it can be
represented in terms of layer potentials. Given the boundary
conditions~\eqref{eqn:jump1} and~\eqref{eqn:jump2} required on the
interface, the fluid velocity can be represented as the single-layer
potential
\begin{align}
  \uu(\xx) = \uu_\infty(\xx) + \SS[\ff_\mathrm{mem}](\xx), \quad
    \xx \in \Omega,
\end{align}
where
\begin{align}
  \SS[\ff](\xx) = \frac{1}{4\pi\mu} \int_{\gamma} \left(
    -\mathbf{I} \log\rho + \frac{\rr \otimes \rr}{\rho^2} \right)
    \ff(\yy) ds_{\yy},
\end{align}
where $\rr = \xx - \yy$ and $\rho = \norm{\rr}$. Then, by the
kinematic equation, the dynamics of the vesicle 
\begin{align}
  \label{eqn:vesVelocity}
  \frac{d\xx}{dt} = \uu_\infty(\xx) + \beta (\ff_\mathrm{mem}\cdot\nn)\nn
  + \SS [\ff_\mathrm{mem}](\xx).
\end{align}
Finally, the inextensibility condition is given by
\begin{align}
  \xx_s\cdot\dot{\xx}_s = 0.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Area Dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where $\Lambda$ is the constant tension that acts as a Lagrange
multiplier to satisfy the length conservation. Note that there is no
pressure Lagrange multiplier since the vesicle's area fluctuates. The
force acing on the membrane is found by taking the variation of
$\mathcal{L}$ with respect to $\gamma$, and by setting the total force
to zero, the steady state shape satisfies
\begin{align}
  \kappa_{ss} + \frac{1}{2}\kappa^3 - \Lambda \kappa = 0.
  \label{eqn:curvODE}
\end{align}
with initial conditions $\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.
Integrating the differential equation~\eqref{eqn:curvODE} with respect
to arclength, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \kappa_0^4 - \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE2}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically using Mathematica. For impermeable vesicles,
it is possible for the curvature to be
periodic~\cite{vee-raj-bir-pur2009}. However, the solution
of~\eqref{eqn:curvODE2} depends on the {\em EllipticF} function, and it
can not be made periodic. Therefore, it the only solution
of~\eqref{eqn:curvODE2} is $\kappa(s) = \kappa_0$, meaning that the
steady state shape is a circle. By the inextensibility condition, the
steady state vesicle has a radius of $L/2\pi$, so its curvature and
tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we can compute the transient
area of the vesicle. Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
\end{align}
The derivative of $A(t)$ includes terms due to the time derivative of
$\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, the stretching term is zero
because of the local inextensibility condition. Therefore,
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential and the divergence theorem.  Finally, using
integration by parts and the Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align}
Finally, using the relationship between the Lagrange multiplier
$\sigma$, and the tension $\Lambda$, the area of a semi-permeable
vesicle is governed by
\begin{align}
  \dot{A}(t) = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that this is consistent with the steady state curvature and
tension~\eqref{eqn:SSshape} since these constant values result in the
integrand in equation~\eqref{eqn:areaROC} being zero.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
We consider a single semi-permeable vesicle in four different flow
regimes. All vesicles have a length of $L=4\pi$, but the area is not
conserved because of semi-permeability. However, the area is bounded
above by the area of a circle with circumference $L=4\pi$. Therefore,
$A(t) \leq 4\pi \approx 12.56$ for all simulations. For each simulation,
we use an adaptive second-order time stepping method~\cite{qua-bir2016}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow}
We start by considering a semi-permeable vesicle in a quiescent flow.
For an impermeable vesicle, the vesicle relaxes to a steady-state shape
with negative and constant tension~\cite{kra-win-sei-lip1996}. However,
from equation~\eqref{eqn:SSshape}, we know that a semipermeable vesicle
will relax to a circle with positive constant tension. The area of a
semi-permeable vesicle is governed by~\eqref{eqn:areaROC}, so we expect
that the time to reach steady state scales with $\beta$. In
Figure~\ref{fig:ellipseArea}, we plot the area of a vesicle with
different permeability rates. The area is computed directly from the
vesicle shape (solid lines) and by integrating
equation~\eqref{eqn:areaROC} (marks) with a first-order method. We also
include black dashed lines indicating the times, $t_{\mathrm{SS}}$, that
the area of the vesicle first exceeds 12.4, meaning that its area is
within 1.3\% of the steady state area.  These times are reported in
Table~\ref{tbl:ellipseRelaxTime}, and we see that the relaxation time
does scale with $\beta$.

\begin{table}[htp]
  \centering
  \begin{tabular}{|C{1cm}|*{6}{C{2cm}}|}
    \hline
    $\beta$ & $10^{0}$ & $10^{-1}$ & $10^{-2}$ & 
             $10^{-3}$ & $10^{-4}$ & $10^{-5}$ \\
    $t_\mathrm{SS}$ & $6.45 \times 10^{0}$ & $2.00 \times 10^{1}$ & 
                      $9.61 \times 10^{1}$ & $8.36 \times 10^{2}$ & 
                      $8.15 \times 10^{3}$ & $8.03 \times 10^{4}$ \\
    ratio & --- & 3.10 & 4.81 & 8.71 & 9.74 & 9.85 \\
    \hline
  \end{tabular}
  \caption{\label{tbl:ellipseRelaxTime} The time, $t_{\mathrm{SS}}$, for
  a semi-permeable vesicle to reach steady state.}
\end{table}

\begin{figure}[htp]
\centering
%\ifTikz
\input{figures/ellipseArea.tikz}
%\fi
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area directly from the vesicle
  shape, while the marks are computed by integrating
  equation~\eqref{eqn:areaROC}. The dashed black lines are the times
  when the vesicle's area is within 1\% of its maximum value of $4\pi$.}
\end{figure}

We further analyze the transient dynamics for small $\beta$ by
considering the contribution of each term in~\eqref{eqn:vesVelocity} to
the vesicle velocity. Initially, the semi-permeability term $\beta(\ff
\cdot \nn)\nn$ is much smaller than the velocity due to the force
balance $\SS[\ff]$. Therefore, the velocity resembles that of an
impermeable vesicle.  However, as the vesicle approaches the impermeable
limit, the velocity contribution from the force balance decays below the
contribution from semi-permeability, and we expect the vesicle to begin
inflating due to permeability. In Figure~\ref{fig:vesVelocity}, we plot
the velocities at three times due to force balance (blue) and
semi-permeability (red), where $\beta = 10^{-3}$. We also plot the
velocity of an identically initialized impermeable vesicle. At $t=0.5$,
the velocity due to semi-permeability is small relative to the velocity
due to force balance, and the semi-permeable vesicle velocity matches
with the impermeable vesicle velocity. At time $t=5$, the impermeable
vesicle is near its steady state shape, and its velocity is comparable
to the semi-permeable velocity, so the vesicle's area begins to
increase. Then, at time $t=50$, the impermeable vesicle is at its steady
state, while the semi-permeable vesicle still has contributions due to
semi-permeability and force balance. In
Figure~\ref{fig:vesVelocityNorm}, we plot the max norm of the different
velocity contributions as a function of time. The black dashed lines
correspond to the times reported in Figure~\ref{fig:vesVelocity}. Again,
we see that the velocity of a semi-permeable vesicle is initially agrees
with the velocity of an impermeable vesicle. Then, as the velocity due
to force balance approaches zero, the semi-permeable term becomes
significant, causing the vesicle to inflate until it reaches a circular
shape.

\begin{figure}[htp]
  \centering
%  \ifTikz
  \input{figures/ellipseVelocity1.tikz}
  \input{figures/ellipseVelocity2.tikz}
  \input{figures/ellipseVelocity3.tikz}
%  \fi
  \caption{\label{fig:vesVelocity} The velocity contributions due to
  semi-permeability (red) and force balance (blue). The green dashed
  line is the velocity of an impermeable vesicle with an identical
  initial shape.}
\end{figure}

\begin{figure}[htp]
%  \ifTikz
  \input{figures/ellipseVelocityNorms.tikz}
%  \fi
  \caption{\label{fig:vesVelocityNorm} The max norm of the velocity due
  to the force balance (blue) and semi-permeability. Also included is
  the velocity of an impermeable vesicle (green). The three black lines
  correspond to the times reported in Figure~\ref{fig:vesVelocity}.}
\end{figure}

\todo[inline]{Finally, we tension and flux of a non-standard shape
vesicle.}

\begin{figure}[htp]
\begin{minipage}{0.40\textwidth}
\ifTikz
\input{figures/starTension.tikz}
\fi
\end{minipage}
\hfill
\begin{minipage}{0.55\textwidth}
  \centering
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime1.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime2.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime3.pdf} \\
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime4.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime5.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime6.pdf}
\end{minipage}
  \caption{\label{fig:starTension} The tension of the time steps shown
  in Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$
  is also included. At this point, the vesicle is nearly circular and
  the tension is constant and positive. The constant is given by
  equation~\eqref{eqn:SSshape}, and this value corresponds to the
  dashed black line.}
\end{figure}

\begin{figure}[htp]
\begin{minipage}{0.40\textwidth}
\ifTikz
\input{figures/starFlux.tikz}
\fi
\end{minipage}
\hfill
\begin{minipage}{0.55\textwidth}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime1.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime2.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime3.pdf} \\
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime4.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime5.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime6.pdf}
\end{minipage}
  \caption{\label{fig:starFlux} The flux of the time steps shown in
  Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$ is
  also included. At this point, the vesicle is nearly circular and there
  is no flux.}
\end{figure}

@@

\ifTikz
\input{figures/shearPhaseDiagram.tikz} 
\fi

@@

When the semi-permeable vesicle reaches a steady-state shape, we have
\begin{align}
  \int_{\gamma} \left(-\frac{\kappa^3}{2} + \kappa \Lambda \right) \, ds
  = 0.
\end{align}
Based on numerical observations, the steady-state shape is a circular
vesicle with constant tension. Since the tension and curvature are
constant, we can solve for the curvature in terms of the tension $\kappa
= \sqrt{2\Lambda}$. Therefore, the steady state shape is a circle of
radius $r = 1/\sqrt{2\Lambda}$, and therefore the length is
\begin{align}
  L = \frac{\sqrt{2}\pi}{\sqrt{\Lambda}},
\end{align}
which is constant by the local inextensibility constraint. Finally, we
can compute the final tension in terms of the length using the formula
\begin{align}
  \Lambda = \frac{2\pi^2}{L^2}.
\end{align}
Therefore, in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

We start by simulating a star-shaped vesicle in a quiescent flow with
semi-permeability rate $\beta=1$ (Figure~\ref{fig:starShape}). To
perform this simulation, we use an adaptive second-order time stepping
method~\cite{qua-bir2016}. The times and reduced areas are in the
titles, and note that the shapes are not eqiuspaced in time.
\begin{figure}[htp]
  \begin{minipage}{0.55\textwidth}
%    \ifTikz
    \input{figures/starRelaxation1.tikz}
    \input{figures/starRelaxation2.tikz}
    \input{figures/starRelaxation3.tikz}
    \input{figures/starRelaxation4.tikz}
    \input{figures/starRelaxation5.tikz}
    \input{figures/starRelaxation6.tikz}
%    \fi
  \end{minipage}
  \hfill
  \begin{minipage}{0.4\textwidth}
%  \ifTikz
  \input{figures/starAreaDeriv.tikz}
%  \fi
  \end{minipage}
  \caption{\label{fig:starShape} A initially star-shaped semi-permeable
  vesicle in a shear quiescent flow with $\beta=1$. Note that the time
  steps are not equispaced. The area (red) and its derivative (blue) of
  an initially star-shaped vesicle as a function of time. The dots
  correspond to the time steps shown in six time steps
  show on the left.}
\end{figure}
